\texttt{Taxonomic supertree construction with}\texttt{\emph{ Incertae sedis}}\texttt{ taxa}

### 2018-06-15 doc/incertae_sedis.tex
very minor

Some taxonomies indicate which taxa are thought to be the result of
hybridization, but the supertree approach described here does not consider any
hybridization. Thus the taxonomic hierarchy can be converted to a tree. While
not all taxonomists believe that named taxa should correspond to clades \citep[see,
for example,][]{HorandlS2010}, the priniciples of phylogenetic classification
are used by such a large majority of practicing taxonomists, that we have chosen
to treat the taxonomic tree that mirrors the hierarchy of the taxonomy as an
estimate of phylogenetic relationships. We refer to labels of taxonomy nodes as
``taxon names'', and assume that taxon names and taxonomy nodes have a
one-to-one correspondence. Note that the taxonomy may also contain out-degree
1 nodes. These nodes correspond to monotypic taxa which contain a single child
of lower taxonomic rank.

@@

\caption{\label{fig:Handling-incertae-sedis}Handling \emph{incertae sedis
}taxa recovers additional edges and taxon names.
Supertree construction on input tree (a) and taxonomy tree (b) with \emph{incertae sedis}
clade $C$ leads to synthesis tree (c) if \emph{incertae sedis} annotations are
ignored, but to synthesis tree (d) if \emph{incertae sedis} annotations are respected.
The tree in (d) recovers taxon names $AB$ and $A$, as well as the edge to clade $AB$
that groups $A$ and $B$. Taxon names and edges that are conditional on handling
\emph{incertae sedis} are shaded grey. \mthcomment{We could use triangle to indicate that $D$ and $B$ could be leaves or clades.}
}

@@

This approach incorrectly treats the intrusion of an \emph{incertae sedis} taxon
into a sibling taxon as a conflicting with monophyly of the sibling taxon. For
example, in Fig. \ref{fig:Handling-incertae-sedis}, the placement of the
\emph{incertae sedis} taxon $C$ within $A$ and $AB$ is considered
to contradict the monophyly of $A$ and $AB$ and thus to conflict with those taxa.
This has three main effects.  These effects are illustrated in the difference
between the synthesis tree in Fig. \ref{fig:Handling-incertae-sedis}c, which
ignores \emph{incertae sedis} annotations, and the synthesis tree in Fig.
\ref{fig:Handling-incertae-sedis}d, which respects them.

>>>>

Some taxonomies indicate which taxa are thought to be the result of
hybridization, but the supertree approach described here does not consider any
hybridization. Thus the taxonomic hierarchy can be converted to a tree. While
not all taxonomists believe that named taxa should correspond to clades \citep[see,
for example,][]{HorandlS2010}, the priniciples of phylogenetic classification
are used by such a large majority of practicing taxonomists.
Thus we have chosen
to treat the taxonomic tree that mirrors the hierarchy of the taxonomy as an
estimate of phylogenetic relationships. We refer to labels of taxonomy nodes as
``taxon names'', and assume that taxon names and taxonomy nodes have a
one-to-one correspondence. Note that the taxonomy may also contain out-degree
1 nodes. These nodes correspond to monotypic taxa which contain a single child
of lower taxonomic rank.

@@

\caption{\label{fig:Handling-incertae-sedis}Handling \emph{incertae sedis
}taxa recovers additional edges and taxon names.
Panel (a) shows the only phylogenetic to the supertree algorithm;
(b) shows taxonomy tree the edge subtending the \emph{incertae sedis}
clade $C$ shown using a dashed line.
The tree in (c) would be the output of the supertree algorithm
if \emph{incertae sedis} annotations are simply ignored.
The tree shown in (d) is the supertree produced 
if \emph{incertae sedis} annotations are respected.
The tree in (d) recovers taxon names $AB$ and $A$, as well as the edge to clade $AB$
that groups $A$ and $B$. Taxon names and edges that are conditional on handling
\emph{incertae sedis} are shaded grey. \mthcomment{We could use triangle to indicate that $D$ and $B$ could be leaves or clades.}
}

@@

This approach incorrectly treats the intrusion of an \emph{incertae sedis} taxon
into a sibling taxon as a conflicting with monophyly of the sibling taxon.
Consider, for example, the trees shown in Fig. \ref{fig:Handling-incertae-sedis}.
In this figure (and throughout this chapter) dashed lines are used for edges that
connect an \emph{incertae sedis} taxon to its parent.
Note that if we were to use the naive semantics,  the placement of the
\emph{incertae sedis} taxon $C$ within $A$ and $AB$ is considered
to contradict the monophyly of $A$ and $AB$ and thus to conflict with those taxa.
This has three main effects.  These effects are illustrated in the difference
between the synthesis tree in Fig. \ref{fig:Handling-incertae-sedis}c, which
ignores \emph{incertae sedis} annotations, and the synthesis tree in Fig.
\ref{fig:Handling-incertae-sedis}d, which respects them.
